### 2020-10-28 soyherb.Rmd
nnn

---
title: "Soybean herbicide study - Dan Smith"
author: "Maxwel Coura Oliveira"
date: "10/28/2020"
output: 
  html_document: 
    toc: yes
    highlight: zenburn
    theme: simplex
    number_sections: yes
---

# Load packages

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(emmeans)
library(glmmTMB)
library(kableExtra)
```

# Load dataset

```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- read_csv("SB12.csv")
```

## Glimpse dataset

```{r echo=TRUE, message=FALSE, warning=FALSE}
glimpse(data)
```

## Skim dataset

```{r echo=TRUE, message=FALSE, warning=FALSE}
skimr::skim(data)
```

# Data visualization

## Waterhemp control at 14 DAT

```{r fig.align='center'}
ggplot(data, aes(x=reorder(other,waterhempcontrol_14), y=waterhempcontrol_14,
       fill=trade, color=trade)) +
  geom_boxplot(color="black") +
  geom_jitter(alpha=0.2) +
  facet_grid(year ~ location) +
  coord_flip() +
  labs(x="", y="Waterhemp control (%)") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Waterhemp control at 28 DAT

```{r fig.align='center'}
ggplot(data, aes(x=reorder(other,waterhempcontrol_28), y=waterhempcontrol_28,
       fill=trade, color=trade)) +
  geom_boxplot(color="black") +
  geom_jitter(alpha=0.2) +
  facet_grid(year ~ location) +
  coord_flip() +
  labs(x="", y="Waterhemp control (%)") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Waterhemp control at harvest

```{r fig.align='center', message=FALSE, warning=FALSE}
ggplot(data, aes(x=reorder(other,waterhempcontrol_harvest), y=waterhempcontrol_harvest,
       fill=trade, color=trade)) +
  geom_boxplot(color="black") +
  geom_jitter(alpha=0.2) +
  facet_grid(year ~ location) +
  coord_flip() +
  labs(x="", y="Waterhemp control (%)") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Yield (bu/acre)

```{r message=FALSE, warning=FALSE, fig.align='center'}
ggplot(data, aes(x=reorder(other,yield_bu), y=yield_bu,
       fill=trade, color=trade)) +
  geom_boxplot(color="black") +
  geom_jitter(alpha=0.2) +
  facet_grid(year ~ location) +
  coord_flip() +
  labs(x="", y="Waterhemp control (%)") +
  theme_minimal() +
  theme(legend.position = "none")
```
# Data wrangling

```{r}
new_dt <- 
  data %>% 
  rename (herbicide = other) %>% 
  mutate(
    wt_14 = waterhempcontrol_14/100,
    wt_28 = waterhempcontrol_28/100,
    yield_kg = yield_bu * 67.5) %>% 
  mutate(
    wt_14 = 
      case_when(
        waterhempcontrol_14 == 100   ~  0.99,
        TRUE            ~ wt_14),
    wt_28 = case_when(
        waterhempcontrol_28 == 100   ~  0.99,
        waterhempcontrol_28 == 0   ~  0.01,
        TRUE            ~ wt_28)) %>% 
  filter(
    herbicide != "PRE only"
  )
```

# Data analysis

## Waterhemp control at 14 DAT

### Model

* Model using herbicide trts, year, and location as fixed effects. Only rep as random effects

```{r message=FALSE, warning=FALSE}
model_14 <- glmmTMB(wt_14 ~ herbicide * location * year + (1|rep), beta_family(link = "logit"), data=new_dt)
```

### Anova

```{r message=FALSE, warning=FALSE}
glmmTMB:::Anova.glmmTMB(model_14)
```

* Anova shows no interaction of year, herbicide and location. There is only herbicide and year effects

### New model

Model using herbicide trt as fixed, and rep, year, and location as random effects

```{r message=FALSE, warning=FALSE}
model_14_2 <- glmmTMB(wt_14 ~ herbicide + (1|rep) + (1|location/year), beta_family(link = "logit"), data=new_dt)
```

### New Anova

```{r message=FALSE, warning=FALSE}
glmmTMB:::Anova.glmmTMB(model_14_2)
```

- No effects of herbicide treatments. However, I am looking into herbicide effects.

### Least square means

```{r }
lsmeans_14<- emmeans(model_14_2, ~ herbicide, cont="pairwise", adjust="none", type="response", alpha=0.05)
```

### Compact letter display

```{r fig.align='center'}
plot(lsmeans_14, ~ herbicide, comparisons=TRUE, type="response", alpha=0.05, adjust="none")
```

```{r message=FALSE, warning =FALSE}
CLD(lsmeans_14$emmeans, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>%
  kbl() %>%
  kable_classic_2(full_width = F)
```

## Waterhemp control at 28 DAT

### Model

* Model using herbicide trts, year, and location as fixed effects. Only rep as random effects

```{r message=FALSE, warning=FALSE}
model_28 <- glmmTMB(wt_28 ~ herbicide * location * year + (1|rep), beta_family(link = "logit"), data=new_dt)
```

### Anova

```{r message=FALSE, warning=FALSE}
glmmTMB:::Anova.glmmTMB(model_28)
```

* Anova shows no interaction of year, herbicide and location. There is only herbicide and year effects

### New model

Model using herbicide trt as fixed, and rep, year, and location as random effects

```{r message=FALSE, warning=FALSE}
model_28_2 <- glmmTMB(wt_28 ~ herbicide + (1|rep) + (1|location/year), beta_family(link = "logit"), data=new_dt)
```

### New Anova

```{r message=FALSE, warning=FALSE}
glmmTMB:::Anova.glmmTMB(model_28_2)
```

- No effects of herbicide treatments

### Least square means

#### Herbicide

```{r }
lsmeans_28<- emmeans(model_28_2, ~ herbicide, cont="pairwise", adjust="none", type="response", alpha=0.05)
```

### Plot

```{r fig.align='center'}
plot(lsmeans_28, ~ herbicide, comparisons=TRUE, type="response", alpha=0.05, adjust="none")
```

### Compact letter display

```{r message=FALSE, warning =FALSE}
CLD(lsmeans_28$emmeans, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>%
  kbl() %>%
  kable_classic_2(full_width = F)
```

```{r include=FALSE}
if(requireNamespace("multcomp")) {
    multcomp::cld(lsmeans_28$emmeans, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE)
}
```