### 2021-10-05 tables.Rmd
as

---
title: "tables"
author: "Maxwel Coura Oliveira"
date: "10/5/2021"
output: html_document
---

```{r}
library(tidyverse)
library(gtExtras)
library(gt)
```

```{r}
tables <- read_csv("tables.csv")
```

```{r}
tables %>%
  mutate_if(is.double, ~round(.,2)) |> 
#  mutate(Parameter = fct_recode(Parameter,
#                                "glufosinate vs glufosinate + <i>S</i>-metolachlor" = "glufosinate vs glufosinate + S-metolachlor")) |> 
  gt() %>% 
  gt_theme_espn() |> 
  tab_spanner(
    label = "Bayes Factor",
    columns = c(`14 DAT`:Yield)) |> 
    tab_footnote(
      footnote = "DAT, Days After Treatment - Waterhemp control",
      locations = cells_column_labels(
      columns = c(`14 DAT`,`28 DAT`)
    )) |> 
  gt_highlight_rows(rows = c(1,3:5,7), 
                    font_weight = "bold", 
                    fill = "beige")  -> fig_tab#|> 
#  gt_hulk_col_numeric(`14 DAT`:Yield, reverse = FALSE, trim = TRUE)

gtsave(fig_tab, "figures/Figure 3.png")
```

```{r}
bayes <- read_csv("bayes1.csv")
```

```{r}
bayes |> 
  gt() %>% 
  gt_theme_espn() -> fig_tab1

gtsave(fig_tab1, "figures/Figure 1.png")
```