### 2021-10-14 Weather.Rmd
asd

---
title: "Weather"
author: "Maxwel Coura Oliveira"
date: "10/7/2021"
output: html_document
---

```{r}
library(tidyverse) # Data wrangling
library(lubridate) # Dates operations
library(daymetr)
library(chirps)
library(patchwork)
library(ggtext)
library(ggthemes)
library(nasapower)
library(ggdist)
library(vegan)
#https://cran.r-project.org/web/packages/ggdist/vignettes/slabinterval.html
```

```{r}
# your data
df_time <- tibble(ID = c('1','2'), 
                  Site = c('Lancaster','Brooklyn'),
                  # Both coordinates in decimal format
                  latitude = c(42.83, 42.52),
                  longitude = c(-90.76, -89.23),
                  Start = c('1989-01-01','1989-01-01'),
                  End = c('2020-12-31','2020-12-31')) %>% 
            mutate_at(5:6, ymd) # as date
```

## Daymet

```{r function}
source("weather_daymet.R")
```

```{r get_data}
#dpp default = 0 - dpp = days prior start date
df_daymet <- weather_daymet(input = df_time, dpp = 0) 
```

```{r}
df_daymet %>% 
  janitor::clean_names() %>% 
  mutate(month_lab = month(date, label = TRUE)) -> weather
```

```{r}
weather %>% 
  filter(month %in% c(5, 6, 7, 8, 9, 10)) %>% 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2019", "2020"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) |> 
  mutate(temp_F = ((tmean * 9/5) + 32)) -> weather1
```

```{r}
weather1 %>% 
#  filter(year %in% c(2019, 2020)) %>% 
  group_by(site, year, month, month_lab) %>% 
  summarise(prec = sum(pp), tmax = mean(tmax), tmin = mean(tmin), tmean = mean(tmean)) %>% 
  mutate_if(is.double, ~round(., 1)) 
```

```{r}
weather1 %>% 
  filter(month %in% c(5, 6, 7, 8, 9, 10)) %>% 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2019", "2020"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) -> weather2
```

```{r}
text1 <- tibble(site = "Brooklyn")
text2 <- tibble(site = "Lancaster")
```

```{r}
weather2 %>% 
  filter(variable != "NA") %>% 
  ggplot(aes(x = month_lab, y = temp_F, 
             color = variable,
             fill = variable,
             group = variable)) +
  geom_jitter(data = weather2 %>% 
                filter(variable == "NA"), show.legend = FALSE, 
              color = "#333333", alpha = 0.02, shape = 16) +
  stat_gradientinterval(position = "dodge", fill_type = "gradient") +
  theme_bw() +
  geom_richtext(data = text1, aes(x = 3.4, y = 44.6), lineheight = 0.4, 
            label = "Mean air temperature <br>in <b style='color:2;'>2019</b> and <b style='color:4;'>2020</b>", size = 2, 
            label.color = NA, fill = NA, inherit.aes = FALSE) +
  geom_richtext(data = text2, aes(x = 3.4, y = 50), lineheight = 0.4, 
            label = "30 years <br>daily mean <br>air temperature", size = 2, 
             label.color = NA, fill = NA, 
            color = "#333333", 
            alpha = 0.5,
            inherit.aes = FALSE) +
  geom_curve(data = text2, aes(x = 3.9, y = 51.8, xend = 4.5, yend = 53.6),
             arrow = arrow(length = unit(0.05, "inch")), 
             size = 0.4, alpha = 0.5, curvature = -0.3, color = "#000000",
             inherit.aes = FALSE) +
  scale_color_manual(values = c(2, 4, "#333333")) +
  scale_fill_manual(values = c(2, 4, "#333333")) +
  facet_grid(~ site) +
  labs(x = NULL, y = "Mean air temperature (C)") +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold", size = 11),
        plot.background = element_rect(fill = "#F9F6EE"),
        panel.background = element_rect(fill = "#F9F6EE")) -> fig1

ggsave("weather.png", height = 3, width = 6)
```

```{r}
weather2 %>% 
  filter(month %in% c(5, 6, 7, 8, 9, 10)) %>% 
  group_by(month_lab, site, year) %>% 
  summarise(pp = sum(pp)) %>% 
  mutate(pp_in = pp / 2.54) |> 
  mutate(year = as.character(year),
         highlight = ifelse(year %in% c("2018", "2019"), TRUE, FALSE),
         variable = if_else(highlight == TRUE, year, "NA")) -> weather2_pp
```

```{r}
weather2_pp |> 
  filter(variable == "NA") |> 
  ggplot(aes(x = month_lab, y = pp_in, fill  = variable)) +
  geom_bar(data = weather2_pp |> filter(variable != "NA"),
           position = position_dodge2(width = 0.4),
           stat = "identity") +
  geom_jitter(alpha = 0.05, position = position_jitter(seed = 242)) +
  scale_fill_manual(values = c(2, 4, "#333333")) +
  theme_bw() +
  labs(x = NULL, y = "Precipitation (in)") +
  geom_richtext(data = text1, aes(x = 3.2, y = 126), lineheight = 0.4, 
            label = "Monthly sum precipitation <br>in <b style='color:red;'>2018</b> and <b style='color:blue;'>2019</b>", size = 2, 
            label.color = NA, fill = NA, inherit.aes = FALSE) +
  geom_richtext(data = text2, aes(x = 3, y = 126), lineheight = 0.4, 
            label = "30 years <br>monthly sum <br>precipitation", size = 2, 
            label.color = NA, 
            fill = NA, 
            alpha = 0.5, 
            color = "#333333",
            inherit.aes = FALSE) +
  geom_curve(data = text2, 
             x = 3.6, y = 125, 
             xend = 3.95, yend = 126,
             alpha = 0.5, 
             arrow = arrow(length = unit(0.05, "inch")), 
             size = 0.4, curvature = -0.3, color = "#333333",
             inherit.aes = FALSE) + 
  facet_grid(~ site) +
  theme(legend.position = "none",
        strip.text = element_markdown(face = "bold", size = 11),
        plot.background = element_rect(fill = "#F9F6EE"),
        panel.background = element_rect(fill = "#F9F6EE")) -> fig2

ggsave("pp_in.png", height = 3, width = 6)
```

```{r}
library(patchwork)
```

```{r}
fig1 / fig2 +
  plot_annotation(tag_levels = 'A')

ggsave("pp_ine.png", height = 6, width = 6)
```